---
title: 'Data Analysis Assignment #1 (50 points total)'
author: "Lee, Ryan"
output:
  html_document: default
---
The following code chunk will:

(a) load the "ggplot2", "gridExtra" and "knitr" packages, assuming each has been installed on your machine,
(b) read-in the abalones dataset, defining a new data frame, "mydata,"
(c) return the structure of that data frame, and
(d) calculate new variables, VOLUME and RATIO.

Do not include package installation code in this document. Packages should be installed via the Console or 'Packages' tab. You will also need to download the abalones.csv from the course site to a known location on your machine. Unless a *file.path()* is specified, R will look to directory where this .Rmd is stored when knitting.

```{r analysis_setup1, message = FALSE, warning = FALSE}

# a) Load the ggplot2 and gridExtra packages.

library(ggplot2)
library(gridExtra)
library(knitr)

# b) Use read.csv() to read the abalones.csv into R, assigning the data frame to "mydata."

mydata <- read.csv("abalones.csv", sep = ",")


# c) Use the str() function to verify the structure of "mydata." You should have 1036 observations
# of eight variables.



# d) Define two new variables, VOLUME and RATIO. Use the following statements to define VOLUME and
# RATIO as variables appended to the data frame "mydata."

mydata$VOLUME <- mydata$LENGTH * mydata$DIAM * mydata$HEIGHT
mydata$RATIO <- mydata$SHUCK / mydata$VOLUME
str(mydata)
```

### Test Items starts from here - There are 6 sections ##########################

##### Section 1: (6 points) Summarizing the data.

(1)(a) (1 point) Use *summary()* to obtain and present descriptive statistics from mydata.  Use table() to present a frequency table using CLASS and RINGS. There should be 115 cells in the table you present.  

```{r Part_1a}
summary(mydata)
mytable <- table(Class = mydata$CLASS,Rings =  mydata$RINGS)
mytable
```

**Question (1 point):  Briefly discuss the variable types and distributional implications such as potential skewness and outliers.**

***Answer: (With right skewness we need to inspect if outliers are affecting the distribution. WHOLE, SHUCK, RINGS, VOLUME and RATIO seem to have large range between min and max.)***

(1)(b) (1 point) Generate a table of counts using SEX and CLASS. Add margins to this table (Hint: There should be 15 cells in this table plus the marginal totals. Apply *table()* first, then pass the table object to *addmargins()* (Kabacoff Section 7.2 pages 144-147)).  Lastly, present a barplot of these data; ignoring the marginal totals.  

```{r Part_1b}
counts <- table(mydata$SEX, mydata$CLASS)
counts.marginal <- addmargins(counts)
counts.marginal
barplot(counts, legend.text = c("Female","Infant","Male"), col = c('pink', 'yellow', 'blue'), beside = TRUE, main = "CLASS membership, Differentiated by SEX", ylab =  'Frequency')


```

**Essay Question (2 points):  Discuss the sex distribution of abalones.  What stands out about the distribution of abalones by CLASS?**

***Answer: (It is interesting that as there are more infants in A2 then any other group, Across all classes we have more males, but see more males then females in A1-3. Meanwhile, A4 and A5 have a more equally split distribution. ))***

(1)(c) (1 point) Select a simple random sample of 200 observations from "mydata" and identify this sample as "work." Use *set.seed(123)* prior to drawing this sample. Do not change the number 123. Note that *sample()* "takes a sample of the specified size from the elements of x." We cannot sample directly from "mydata." Instead, we need to sample from the integers, 1 to 1036, representing the rows of "mydata." Then, select those rows from the data frame (Kabacoff Section 4.10.5 page 87). 

Using "work", construct a scatterplot matrix of variables 2-6 with *plot(work[, 2:6])* (these are the continuous variables excluding VOLUME and RATIO). The sample "work" will not be used in the remainder of the assignment. 

```{r Part_1c, fig.height=6, fig.width=6}
set.seed(123)
work <- mydata[sample(nrow(mydata), 200, replace = FALSE),]
plot(work[, 2:6])
```

-----

##### Section 2: (5 points) Summarizing the data using graphics.

(2)(a) (1 point) Use "mydata" to plot WHOLE versus VOLUME. Color code data points by CLASS.

```{r Part_2a}

ggplot(data = mydata, aes(x=mydata$VOLUME, y=mydata$WHOLE, color = mydata$CLASS,)) + 
  geom_point() +
  xlab("Volume")+
  ylab("WHole") +
  labs(title = "Whole vs. Volume ", col="Class")

```


(2)(b) (2 points) Use "mydata" to plot SHUCK versus WHOLE with WHOLE on the horizontal axis. Color code data points by CLASS. As an aid to interpretation, determine the maximum value of the ratio of SHUCK to WHOLE.  Add to the chart a straight line with zero intercept using this maximum value as the slope of the line. If you are using the 'base R' *plot()* function, you may use *abline()* to add this line to the plot. Use *help(abline)* in R to determine the coding for the slope and intercept arguments in the functions. If you are using ggplot2 for visualizations, *geom_abline()* should be used.

```{r Part_2b}

max_value <- max(mydata$SHUCK / mydata$WHOLE)


ggplot(data = mydata, aes(x=mydata$WHOLE, y=mydata$SHUCK, color = mydata$CLASS,)) + 
  geom_point() +
  xlab("WHOLE (g)")+
  ylab("SHUCK (g)") +
  geom_abline(slope = max_value)

```

**Essay Question (2 points):  How does the variability in this plot differ from the plot in (a)?  Compare the two displays.  Keep in mind that SHUCK is a part of WHOLE.  Consider the location of the different age classes.**

***Answer: (In the second plot there is less noise. Also we see that as class increases we see more noise. But as class increases the shuck weight gets more disperse, perhaps shuck is dispersed by something other than whole weight. Further analysis might give insight. A3 seems to be the most correlated linearly perhaps older abalones do not add additoonal value as they grow older.)***

-----

##### Section 3: (8 points) Getting insights about the data using graphs.

(3)(a) (2 points) Use "mydata" to create a multi-figured plot with histograms, boxplots and Q-Q plots of RATIO differentiated by sex. This can be done using *par(mfrow = c(3,3))* and base R or *grid.arrange()* and ggplot2. The first row would show the histograms, the second row the boxplots and the third row the Q-Q plots. Be sure these displays are legible.  

```{r Part_3a, fig.height=6, fig.width=6}
sex.I <- mydata$RATIO[mydata$SEX == "I"]
sex.F <- mydata$RATIO[mydata$SEX == "F"]
sex.M <- mydata$RATIO[mydata$SEX == "M"]


par(mfrow = c(3,3))
hist(sex.I, col = "yellow", main = "Infant Ratio", xlab = "")
hist(sex.F, col = "Pink", main = "Female Ratio", xlab = "")
hist(sex.M, col = "Blue", main = "Male Ratio", xlab = "")
boxplot(sex.I, col = "yellow", main = "Infant Boxplot", xlab = "")
boxplot(sex.F, col = "Pink", main = "Female Boxplot", xlab = "")
boxplot(sex.M, col = "Blue", main = "Male Boxplot", xlab = "")
qqnorm(sex.I, col = "yellow", main = "Infant Ratio")
qqline(sex.I)
qqnorm(sex.F, col = "pink", main = "Female QQPlot")
qqline(sex.F)
qqnorm(sex.M, col = "blue", main = "Male QQPlot")
qqline(sex.M)

```

**Essay Question (2 points): Compare the displays.  How do the distributions compare to normality? Take into account the criteria discussed in the sync sessions to evaluate non-normality.**

***Answer: (The QQ plots drift from normal leading to a conclusion that they are not normal. All of the distributions are right skewed for sex. The QQ plots suggest there is that at as we approach 3 quantiles more outliers are occuring which could be skwing the data. Infant has the most outliers as witnessed in the boxplot.)***

(3)(b) (2 points) Use the boxplots to identify RATIO outliers (mild and extreme both) for each sex. Present the abalones with these outlying RATIO values along with their associated variables in "mydata" (Hint:  display the observations by passing a data frame to the kable() function).  

```{r Part_3b}


infants.outliers <- boxplot.stats(sex.I, coef = 1.5)$out
female.outliers <- boxplot.stats(sex.F, coef = 1.5)$out
male.outliers <- boxplot.stats(sex.M, coef = 1.5)$out

I.DF.Out <- subset(mydata, mydata$SEX == "I" & mydata$RATIO %in% infants.outliers)
kable(I.DF.Out)
F.DF.Out <- subset(mydata, mydata$SEX == "F" & mydata$RATIO %in% female.outliers)
kable(F.DF.Out)
M.DF.Out <- subset(mydata, mydata$SEX == "M" & mydata$RATIO %in% male.outliers)
kable(M.DF.Out)

# oer group should be 19 (sex subset)
# use kable, no need to reference previous, boxplot stats
# have a data frame for infants and females and males
```

**Essay Question (2 points):  What are your observations regarding the results in (3)(b)?**

***Answer: (The Infant class has more outliers than the other sexes. Al three categories have more larger outliers than small. Male and Female abalones seem to have most outliers in A3 when compared with ratio. )***

-----

##### Section 4: (8 points) Getting insights about possible predictors.

(4)(a) (3 points) With "mydata," display side-by-side boxplots for VOLUME and WHOLE, each differentiated by CLASS There should be five boxes for VOLUME and five for WHOLE. Also, display side-by-side scatterplots:  VOLUME and WHOLE versus RINGS. Present these four figures in one graphic:  the boxplots in one row and the scatterplots in a second row. Base R or ggplot2 may be used.


```{r Part_4a}


box1 <- ggplot(mydata) + geom_boxplot(aes(x = CLASS, y = VOLUME, color = CLASS)) + labs(x = "Class", y= "Volume")
box2 <- ggplot(mydata) + geom_boxplot(aes(x = CLASS, y = WHOLE, color  = CLASS)) + labs(x = "Class", y= "Whole")
scatter1 <- ggplot(mydata) + geom_point(aes(x = RINGS, y = VOLUME, color = CLASS)) + labs(x = "Rings")
scatter2 <- ggplot(mydata) + geom_point(aes(x = RINGS, y = WHOLE, color = CLASS)) + labs(x = "Rings")

grid.arrange(box1,box2,scatter1,scatter2, layout_matrix = rbind(c(1,2),c(3,4)))

```

**Essay Question (5 points) How well do you think these variables would perform as predictors of age?  Explain.**

***Answer: (I think that you can identify the age of abalone through A3 but A4 and A5 are not detetable by Volume or Whole as shown by the boxplots. Whixh we also see in the scatterplot where rings are clearly delineated as the measure of age.)***

-----

##### Section 5: (12 points) Getting insights regarding different groups in the data.

(5)(a) (2 points) Use *aggregate()* with "mydata" to compute the mean values of VOLUME, SHUCK and RATIO for each combination of SEX and CLASS. Then, using *matrix()*, create matrices of the mean values. Using the "dimnames" argument within *matrix()* or the *rownames()* and *colnames()* functions on the matrices, label the rows by SEX and columns by CLASS. Present the three matrices (Kabacoff Section 5.6.2, p. 110-111). The *kable()* function is useful for this purpose.  You do not need to be concerned with the number of digits presented.

```{r Part_5a}
Vol.Agg <-aggregate(VOLUME~ SEX + CLASS, data = mydata, mean)
Shuck.Agg <- aggregate(SHUCK ~ SEX + CLASS, data = mydata, mean)
Ratio.Agg <- aggregate(RATIO ~ SEX + CLASS, data = mydata, mean)

rowname = c("Female", "Infant", "Male")
colname = c("A1", "A2", "A3", "A4", "A5")

Volume.Matrix <- matrix(round(Vol.Agg$VOLUME, digits = 2), nrow = 3, ncol = 5, dimnames = list(rowname,colname))
Shuck.Matrix <- matrix(round(Shuck.Agg$SHUCK, digits = 2), nrow = 3, ncol = 5, dimnames = list(rowname,colname))
Ratio.Matrix <- matrix(round(Ratio.Agg$RATIO, digits = 4), nrow = 3, ncol = 5, dimnames = list(rowname,colname))


kable(Volume.Matrix)
kable(Shuck.Matrix)
kable(Ratio.Matrix)

```

(5)(b) (3 points) Present three graphs. Each graph should include three lines, one for each sex. The first should show mean RATIO versus CLASS; the second, mean VOLUME versus CLASS; the third, mean SHUCK versus CLASS. This may be done with the 'base R' *interaction.plot()* function or with ggplot2 using *grid.arrange()*.

```{r Part_5b, fig.width = 11, fig.height=4}
overview <- cbind(Vol.Agg,Shuck.Agg$SHUCK,Ratio.Agg$RATIO)
colnames(overview) <- c("Sex", "Class", "Volume", "Shuck", "Ratio")
overviewdf<- data.frame(overview)



lp1 <- ggplot(overviewdf, aes(x = Class, y = Ratio,group = Sex, colour = Sex)) + geom_line() + geom_point(size = 2)+ labs(title = "Ratio per Class")
lp2 <- ggplot(overviewdf, aes(x = Class, y = Volume,group = Sex, colour = Sex)) +geom_point(size = 2)+ geom_line()+ labs(title = "Volume per Class")
lp3 <- ggplot(overviewdf, aes(x = Class, y = Shuck,group = Sex, colour = Sex)) +geom_point(size = 2)+ geom_line() + labs(title = "Shuck per Class")

grid.arrange(lp1,lp2,lp3,layout_matrix = rbind(c(1,2,3),c(1,2,3)))
```

**Essay Question (2 points):  What questions do these plots raise?  Consider aging and sex differences.**

***Answer: (As Volume and Shuch increases there is not way to determine if there is a difference in age for both sexes. As Ratio decreases afe is easier to detemine but why doesn't the ratio change during A1 and A2? )***

5(c) (3 points) Present four boxplots using *par(mfrow = c(2, 2)* or *grid.arrange()*. The first line should show VOLUME by RINGS for the infants and, separately, for the adult; factor levels "M" and "F," combined. The second line should show WHOLE by RINGS for the infants and, separately, for the adults. Since the data are sparse beyond 15 rings, limit the displays to less than 16 rings. One way to accomplish this is to generate a new data set using subset() to select RINGS < 16.  Use ylim = c(0, 1100) for VOLUME and ylim = c(0, 400) for WHOLE.  If you wish to reorder the displays for presentation purposes or use ggplot2 go ahead.

```{r Part_5c}

par(mfrow=c(2,2))
Rings.Subset <- subset(mydata, mydata$RINGS <16)
boxplot(Rings.Subset$VOLUME[Rings.Subset$SEX == "I"] ~ Rings.Subset$RINGS[Rings.Subset$SEX == "I"],main = "Infant Volume by Rings", col = "Yellow", xlab = "Rings", ylab = "Volume", ylim = c(0,1100))
boxplot(Rings.Subset$VOLUME[Rings.Subset$SEX == "F" | Rings.Subset$SEX == "M"] ~ Rings.Subset$RINGS[Rings.Subset$SEX == "F" | Rings.Subset$SEX == "M"], main = "Adult Volume by Rings", col = "purple", xlab = "Rings", ylab = "Volume", ylim = c(0,1100))
boxplot(Rings.Subset$WHOLE[Rings.Subset$SEX  == "I"] ~ Rings.Subset$RINGS[Rings.Subset$SEX == "I"], main = "Infant Whole by Rings", col = "yellow", xlab = "Rings", ylab = "Whole Weight", ylim = c(0,400 ))
boxplot(Rings.Subset$WHOLE[Rings.Subset$SEX  == "F" | Rings.Subset$SEX == "M"] ~ Rings.Subset$RINGS[Rings.Subset$SEX == "F" | Rings.Subset$SEX == "M"], main = " Adult Whole by Rings", col = "purple", xlab = "Rings", ylab = "Whole Weight",  ylim = c(0,400 ))
```

**Essay Question (2 points):  What do these displays suggest about abalone growth?  Also, compare the infant and adult displays.  What differences stand out?**

***Answer: (Abalone volume/whole levels out as they age as measured by rings. But the abalones are more disperses as they age as shown by the boxplot whiskers.)***

-----

##### Section 6: (11 points) Conclusions from the Exploratory Data Analysis (EDA).

**Conclusions**


**Essay Question 1) (5 points)	Based solely on these data, what are plausible statistical reasons that explain the failure of the original study? Consider to what extent physical measurements may be used for age prediction.**

***Answer: (The outliers in this dataset make it difficult to determine a result but it seems as though these measurements of the abalone are not great for determining age as they stay to similar across older age groups. Also is the determination of the sexes mostly around infants accurate because of the high level of outliers and the data making less sense like the ring count in the outlier table for infant.)***

**Essay Question 2) (3 points) Do not refer to the abalone data or study.  If you were presented with an overall histogram and summary statistics from a sample of some population or phenomenon and no other information, what questions might you ask before accepting them as representative of the sampled population or phenomenon?**

***Answer: (How was the data measured? Is this all of the data? How many samples per year? Are all the abalones from one area?  )***

**Essay Question 3)  (3 points)	Do not refer to the abalone data or study.  What do you see as difficulties analyzing data derived from observational studies? Can causality be determined?  What might be learned from such studies?**

***Answer: (Observational studies are prone to higher error as they are being detemined from noiser sources. Identifcation even by a export leads to a higher error rate and can skew the data and prevent conclusions from being made. Causality can potentially be determined from oberservational studies, but that is difficult as they're usually not randomized and and can be affected by outside factors, so it is hard to determine if the causality is from a bias in the data or real. We can learn that these studies are difficult to do and need a lot of randomization and correct identification. Additonally theese studies need to have a control group to compare the data against otherwise there is no comparitor to detect bias and if the data is skewing from normal.)***
